Article figures

### 2021-06-07 src/article_figures.Rmd
Show also atot, q, TPR, and FPR in a table.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
library(knitr)
library(tidyverse)
save_figs = TRUE
fig_path="~/FRCBS/results-for-eba-paper/pdf"
```

@@

  rrfFit.varimp <- rrfFit.varimp %>% 
    mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)),
           gender=factor("both", levels=c("female", "both", "male")))
  rrfFit.varimp <- left_join(combined_variables, rrfFit.varimp, by="Pretty")
  return(rrfFit.varimp)  
}

# Creates a forest plot of means and confidence intervals from posterior distributions.
# Puts both male and female results in the same plot.
create_double_forest_plot <- function(male_posterior, female_posterior, variables, combined_variables, base_size = 11) {
  result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
  cis <- result

@@

get_name_of_data <- function(s) {
  s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
  s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
  # s <- map(s, function(x) when(x, 
  #                              x == "progesa" ~ "eProgesa",
  #                              x == "finngen" ~ "Biobank",
  #                              x == "findonor" ~ "FinDonor",
  #                              ~ "koe"))
  return(s)
}

@@

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model", fill="Deferral length") + 
    scale_y_discrete(labels=get_name_of_data) +
    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "dummy"),
                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Dummy"),
                       values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("variable", scales="free_x") +
    #theme(legend.position = "right", legend.direction = "vertical") 
    theme(legend.position = "bottom", legend.direction = "vertical")
  return(list(g=g, df=df))
}

@@

create_classification_scatter_plot <- function(df, hb_threshold, probability_of_deferral_threshold) {
  xymin <- min(min(df$predicted), min(df$observed))
  xymax <- max(max(df$predicted), max(df$observed))
  if (is.null(probability_of_deferral_threshold)) { # predict deferral based on Hb
    df <- df %>% 
      mutate(new_predicted_label= ifelse(predicted < hb_threshold, 1, 0))
  } else { # predict deferral based on probability of deferral
    df <- df %>% 
      mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0))
  }
  df <- df %>%
    mutate(confusion_class = factor(ifelse(deferral == 1, 
                                    ifelse(new_predicted_label == 1, "True positive", "False negative"),
                                    ifelse(new_predicted_label == 1, "False positive", "True negative")),
           levels=c("True positive", "False negative", "False positive", "True negative")))
  scatter_plot <- ggplot(df, aes(x = observed, y=predicted, color = confusion_class)) +
    geom_point() +
    #xlim(xymin,xymax) + ylim(xymin,xymax) +
    scale_x_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_y_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_colour_discrete(drop=FALSE) +
    geom_abline(intercept = 0, slope = 1) +
    labs(x = "Observed", y = "Predicted", colour = "Deferral status") +
    #scale_colour_discrete(labels=c("Accepted", "Deferred")) +
    geom_smooth(mapping=aes(x = observed, y=predicted), colour="black", show.legend = FALSE) +
    geom_vline(xintercept = hb_threshold, linetype = "dashed") +
    geom_hline(yintercept = hb_threshold, linetype = "dashed") +
    theme(legend.position = "bottom")
    #ggtitle("Observed vs predicted Hb-values")
  return(scatter_plot)
}
```

@@

```{r}
g1 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, NULL)
g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.1)
if (save_figs) {
  ggsave(filename="classification-scatter-plot-hb-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g1, dpi=600, units="mm", width=180)
  ggsave(filename="classification-scatter-plot-probability-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g2, dpi=600, units="mm", width=180)
}
g1
g2
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
library(knitr)
library(tidyverse)
save_figs = FALSE
fig_path="~/FRCBS/results-for-eba-paper/pdf"
```

@@

  rrfFit.varimp <- rrfFit.varimp %>% 
    mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)),
           gender=factor("both", levels=c("female", "both", "male")))
  rrfFit.varimp <- left_join(combined_variables, rrfFit.varimp, by="Pretty")
  return(rrfFit.varimp)  
}
```

```{r, eval=FALSE}
# Creates a forest plot of means and confidence intervals from posterior distributions.
# Puts both male and female results in the same plot.
create_double_forest_plot <- function(male_posterior, female_posterior, variables, combined_variables, base_size = 11) {
  result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
  cis <- result

@@

get_name_of_data <- function(s) {
  s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
  # Note that recode stupidly uses order "old"="new"
  s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
  # s <- map(s, function(x) when(x, 
  #                              x == "progesa" ~ "eProgesa",
  #                              x == "finngen" ~ "Biobank",
  #                              x == "findonor" ~ "FinDonor",
  #                              ~ "koe"))
  return(s)
}

@@

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and model", colour="Sex", shape="Model", fill="Deferral length") + 
    scale_y_discrete(labels=get_name_of_data) +
    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "dummy"),
                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Baseline"),
                       values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("variable", scales="free_x") +
    #theme(legend.position = "right", legend.direction = "vertical") 
    theme(legend.position = "bottom", legend.direction = "vertical")
  return(list(g=g, df=df))
}

@@

create_classification_scatter_plot <- function(df, hb_threshold, probability_of_deferral_threshold) {
  xymin <- min(min(df$predicted), min(df$observed))
  xymax <- max(max(df$predicted), max(df$observed))
  if (is.null(probability_of_deferral_threshold)) { # predict deferral based on Hb
    df <- df %>% 
      mutate(new_predicted_label= ifelse(predicted < hb_threshold, 1, 0))
    prediction_threshold <- hb_threshold
  } else { # predict deferral based on probability of deferral
    df <- df %>% 
      mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0))
    prediction_threshold <- min(df %>% filter(scores <= probability_of_deferral_threshold) %>% pull(predicted))
  }

  cat(sprintf("Prediction threshold is %.1f\n", prediction_threshold))

  # Get the R2 correlation
  r2 <- summary(lm(predicted ~ observed, data=df))$r.squared
  r2_text <- sprintf("%.2f", r2)

  df <- df %>%
    mutate(confusion_class = factor(ifelse(deferral == 1, 
                                    ifelse(new_predicted_label == 1, "True positive", "False negative"),
                                    ifelse(new_predicted_label == 1, "False positive", "True negative")),
           levels=c("True positive", "False negative", "False positive", "True negative")))

  n <- nrow(df)
  fp <- sum(df$confusion_class=="False positive")
  tp <- sum(df$confusion_class=="True positive")
  deferrals <- nrow(df %>% filter(confusion_class %in% c("True positive", "False negative")))
  accepted <- n - deferrals

  cat(sprintf("Incorrectly predicting %.1f%% of accepted donations as deferrals\n", 100*fp/accepted))
  cat(sprintf("Correctly predicting %.1f%% of deferred donations as deferrals\n", 100*tp/deferrals))

  scatter_plot <- ggplot(df, aes(x = observed, y=predicted, color = confusion_class)) +
    geom_point() +
    #xlim(xymin,xymax) + ylim(xymin,xymax) +
    scale_x_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_y_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_colour_discrete(drop=FALSE) +
    geom_abline(intercept = 0, slope = 1) +
    labs(x = "Observed", y = "Predicted", colour = "Deferral status") +
    #scale_colour_discrete(labels=c("Accepted", "Deferred")) +
    geom_smooth(mapping=aes(x = observed, y=predicted), colour="black", show.legend = FALSE) +
    geom_vline(xintercept = hb_threshold, linetype = "dashed") +
    geom_hline(yintercept = prediction_threshold, linetype = "dashed") +
    annotate("text", x=120, y=190, color="black", label=bquote(R^2==.(r2_text)), fontface="bold") +
    theme(legend.position = "bottom")
    #ggtitle("Observed vs predicted Hb-values")
  return(scatter_plot)
}
```

@@

```{r}
g1 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, NULL)
g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.04)
if (save_figs) {
  ggsave(filename="classification-scatter-plot-hb-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g1, dpi=600, units="mm", width=180)
  ggsave(filename="classification-scatter-plot-probability-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g2, dpi=600, units="mm", width=180)
}
g1
g2
```

@@

Create a table of atot and TPR and FPR values for each method/data. Q is the fraction of avoided deferrals.

```{r}
#data_frames <- map(dummy2_ids, get_data_frame)
df2 <- df %>% select(Id, threshold6, threshold12)  # Get the optimal thresholds
```

```{r}
helper <- function(Id, threshold6, threshold12) {
  df2 <- get_data_frame(Id)
  r6 <- get_rates(df2, threshold6)
  r12 <- get_rates(df2, threshold12)
  df <- get_cost(r6$TPR, r6$FPR, r12$TPR, r12$FPR, p = parameters)
  #df <- tibble(a=1, b=2)
  df$Id <- Id
  return(df %>% select(Id, everything()))  
}

d <- get_mikkos_cost_constants()$d  # Fraction of deferrals of all donors
result <- pmap_dfr(df2, helper) 
result <- result %>%
  mutate(across(c(q6, q12), function(x) x / d)) %>%
  rename(Q6=q6, Q12=q12)
kable(result, digits=3)
```
